Authentication based on DRAM PUFs

### 2021-06-17 latexPaper/paper.tex
4 pages

@@

% Write abstract here
\begin{abstract}
Security concerns relating to IoT are increasing and the most prominent of these concerns is that of counterfeit or tampered hardware. For such a problem, it is important to employ a hardware-intrinsic security mechanism for authentication of the device. While previous solutions like secure key storage attempt to solve this problem they only pass on the issue to the next stage. This paper outlines the use of Physically Unclonable Functions (PUFs) for hardware-intrinsic authentication. It summarizes well-known works on DRAM based PUFs and also presents known protocols and schemes to achieve authentication on embedded devices using PUFs.
\end{abstract}

% Each section begins with a \section{title} command
\section{Introduction}
% \IEEEPARstart{}{} creates a tall first letter for this first paragraph
\IEEEPARstart{I}{n} In light of the IoT era the world has seen a spike in the number of smart devices deployed in the market. These ubiquitous smart devices can be found anywhere, from toasters to cars. While they bring a lot of advantages to the table they add a profound amount of concern regarding security and privacy. \cite{EMVT12} The most prominent security concern includes counterfeit or tampered hardware. \cite{PEC06}

@@

PUFs by definition are unclonable and also because they exploit the process variations of manufacturing to form a hardware-based fingerprint. This hardware based fingerprint with a supporting lightweight protocol can then be used for authentication. Clearly, it is very important for the PUF instance to be intrinsic to the hardware that needs to be authenticated. The intrinsic nature can be assured either by addition of dedicated circuits or by using PUFs inherent to the device. Since the latter adds costs to the manufacturing process, inherent and intrinsic PUFs are much preferred. DRAM based PUFs are such a class of PUFs, because they can be found in almost all COTS devices.

This paper briefly touches on the basics of PUFs and dives into how DRAMs are used as PUF instances in the Theory section. Through means of survey it presents different types of authentication mechanisms used in conjunction with DRAM based PUFs. Finally, it also presents the main implementation objectives accompanied by a summary of metrics by means of survey.

\section{Theory}
This sections goes into the basics of PUFs and workings of DRAM PUFs.

\subsection{PUFs}
In practice most PUFs are characterized by challenges and responses. A challenge is the input given to the PUF, which then utilizes the random manufacturing process variations to convert the challenge to a response. The randomness makes the response unpredictable and because these variations cannot be controlled, even by the manufacturer, the response is unclonable on any other device. Hence, the response can be used as a hardware intrinsic fingerprint for authentication.

\begin{figure}[!hbt]
    % Center the figure.
    \begin{center}
    % Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
    \includegraphics[scale=0.5]{figs/PUF.png}
    % Create a subtitle for the figure.
    \caption{PUFS}
    % Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
    \label{fig:tf_plot}
    \end{center}
\end{figure}

@@

\subsection{DRAM PUFs}
There are different methods in which the process variations of DRAM can be leveraged and they are explained below:

@@

% This is how you define a table: the [!hbt] means that LaTeX is forced (by the !) to place the table exactly here (by h), or if that doesnt work because of a pagebreak or so, it tries to place the table to the bottom of the page (by b) or the top (by t).
\begin{table}[!hbt]
    % Center the table
    \begin{center}
    % Title of the table
    \caption{Uniqueness Analysis, Refresh Pause Method}
    \label{tab:simParameters}
    % Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
    \begin{tabular}{|c|c|c|}
        % To create a horizontal line, type \hline
        \hline
        % To end a column type &
        % For a linebreak type \\

        Avg. HD = 216 & Min. HD = 113 & Different module at t=90s \\
        \hline
        Avg. HD = 503 & Min. HD = 486 & Same module at t=60s \\
        \hline

    \end{tabular}
    \end{center}
\end{table}

@@

\textit{Run-time Access.} Based on the methods DRAM PUFs are limited in access. For run-time access of DRAM PUFs it is important that the PUF operation doesn’t hinder the functioning of the OS and applications. The solution presented by Tehranipoor et al restricts the access to the DRAM PUF to an early boot stage because DRAM startup values cannot be accessed later. Whereas, write failures based DRAM PUFs offer the flexibility of run time access.

Refresh pause based  DRAM PUFs can offer run-time access but with modifications to the kernel in some cases. When LPDDR DRAM is being used the PUFs can be accessed at run-time due to “Partial Array Self-Refresh” functionality. \cite{SUT18} When DRAMs are split by different memory controllers they can still provide run-time access by splitting the memory per controller for OS and PUF respectively. Special case arises when the DRAM being used has none of the above features, in such case, memory ballooning is used to reserve a DRAM region and selective memory refreshing is used to refresh selected regions by initiating a read.\cite{XIO16}

@@

\textit{One Way Authentication.} The flowchart in Fig.\ref{fig:one_way} depicts the one-way authentication process

@@

    \bibitem{SUT18}
    Soubhagya Sutar, Arnab Raha, Devadatta Kulkarni, Rajeev Shorey, Jeffrey Tew, and Vijay Raghunathan. 2017. D-PUF: An Intrinsically Reconfigurable DRAM PUF for Device Authentication and Random Number Generation. ACM Trans. Embed. Comput. Syst. 17, 1, Article 17 (January 2018), 31 pages. DOI:https://doi.org/10.1145/3105915

@@

    \bibitem{TEH15}
    Fatemeh Tehranipoor, Nima Karimian, Kan Xiao, and John Chandy. 2015. DRAM based Intrinsic Physical Unclonable Functions for System Level Security. In <i>Proceedings of the 25th edition on Great Lakes Symposium on VLSI</i> (<i>GLSVLSI '15</i>). Association for Computing Machinery, New York, NY, USA, 15–20. DOI:https://doi.org/10.1145/2742060.2742069

    \bibitem{KELLER}
    C. Keller, F. Gürkaynak, H. Kaeslin and N. Felber, "Dynamic memory-based physically unclonable function for the generation of unique identifiers and true random numbers," 2014 IEEE International Symposium on Circuits and Systems (ISCAS), 2014, pp. 2740-2743, doi: 10.1109/ISCAS.2014.6865740.

@@

    \bibitem{HASH}
    M. S. Hashemian, B. Singh, F. Wolff, D. Weyer, S. Clay and C. Papachristou, "A robust authentication methodology using physically unclonable functions in DRAM arrays," 2015 Design, Automation \& Test in Europe Conference \& Exhibition (DATE), 2015, pp. 647-652, doi: 10.7873/DATE.2015.0308.

>>>>

\usepackage[belowskip=-5pt,aboveskip=5pt]{caption}
\setlength{\intextsep}{10pt plus 2pt minus 2pt}

@@

% Write abstract here
\begin{abstract}
One of the major security concerns regarding IoT is that of counterfeit or tampered hardware. To overcome this problem it is important to employ a hardware-intrinsic security mechanism for authentication of the device. While previous solutions like secure key storage attempt to solve this problem they only pass on the issue to the next stage. This paper outlines the use of Physically Unclonable Functions (PUFs) for hardware-intrinsic authentication. It summarizes well-known works on DRAM based PUFs and also presents known protocols and schemes to achieve authentication on embedded devices using PUFs.
\end{abstract}

% Each section begins with a \section{title} command
\section{Introduction}
% \IEEEPARstart{}{} creates a tall first letter for this first paragraph
\IEEEPARstart{I}{n} light of the IoT era the world has seen a spike in the number of smart devices deployed in the market. These ubiquitous smart devices can be found anywhere, from toasters to cars. While they bring a lot of advantages to the table they add a profound amount of concern regarding security and privacy. \cite{EMVT12} The most prominent security concern includes counterfeit or tampered hardware. \cite{PEC06}

@@

The next section exlpains the basics of PUFs and why they are a better solution to the problem. It also dives into how DRAMs can be used as PUFs and explains various methods of doing so.

\section{Theory}
This sections goes into the basics of PUFs and workings of DRAM PUFs.
\vspace{-0.9em}
\subsection{PUFs}

In practice most PUFs are characterized by challenges and responses. A challenge is the input given to the PUF. The PUF utilizes random manufacturing variations to convert the challenge to a response. The randomness makes the response unpredictabl. And because these variations cannot be controlled, even by the manufacturer, the response is unclonable on any other device. Thios way they attain their property of being \textbf{physically unclonable}.

\begin{figure}[!hbt]
    % Center the figure.
    \centering
    % Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
    \includegraphics[scale=0.5]{figs/PUF.png}
    % Create a subtitle for the figure.
    \setlength{\belowcaptionskip}{-10pt}
    \caption{PUFS}
    % Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
    \label{fig:tf_plot}

\end{figure}

@@

\vspace{-0.9em}

\subsection{DRAM PUFs}
There are different methods in which the process variations of DRAM can be leveraged to implement PUFs. The well-known methods are explained below.

@@

% % This is how you define a table: the [!hbt] means that LaTeX is forced (by the !) to place the table exactly here (by h), or if that doesnt work because of a pagebreak or so, it tries to place the table to the bottom of the page (by b) or the top (by t).
% \begin{table}[!hbt]
%     % Center the table
%     \begin{center}
%     % Title of the table
%     \caption{Uniqueness Analysis, Refresh Pause Method}
%     \label{tab:simParameters}
%     % Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
%     \begin{tabular}{|c|c|c|}
%         % To create a horizontal line, type \hline
%         \hline
%         % To end a column type &
%         % For a linebreak type \\

%         Avg. HD = 216 & Min. HD = 113 & Different module at t=90s \\
%         \hline
%         Avg. HD = 503 & Min. HD = 486 & Same module at t=60s \\
%         \hline

%     \end{tabular}
%     \end{center}
% \end{table}

@@

\textit{Run-time Access.} Depending on the method used, DRAM PUFs have restricted access. For run-time access of DRAM PUFs it is important that the PUF operation doesn’t hinder the functioning of the OS and applications. The solution presented by Tehranipoor et al restricts the access to the DRAM PUF to an early boot stage because DRAM startup values cannot be accessed later. Whereas, write failures based DRAM PUFs offer the flexibility of run time access.

Refresh pause based  DRAM PUFs can offer run-time access but with modifications to the kernel in some cases. When LPDDR DRAM is being used the PUFs can be accessed at run-time due to “Partial Array Self-Refresh” functionality. \cite{SUT18} When DRAMs are split by different memory controllers they can still provide run-time access by splitting the memory per controller for OS and PUF respectively. Special case arises when the DRAM being used has none of the above features, in such case, \textit{memory ballooning} is used to reserve a DRAM region and \textit{selective memory refreshing} is used to refresh selected regions by initiating a read.\cite{XIO16}

@@

\textit{One Way Authentication.} The flowchart in Fig.\ref{fig:one_way} depicts the one-way authentication process. In short, The server matches the device's remeasured response against the CRP database. If matches, the device is authenticated

@@

    \bibitem{SUT18}
    Soubhagya Sutar et al. D-PUF: An Intrinsically Reconfigurable DRAM PUF for Device Authentication and Random Number Generation. ACM Trans. Embed. Comput. Syst. 17, 1, Article 17 (January 2018), 31 pages. DOI:https://doi.org/10.1145/3105915

@@

    \bibitem{TEH15}
    Fatemeh Tehranipoor et al. 2015. DRAM based Intrinsic Physical Unclonable Functions for System Level Security. 25th edition on Great Lakes Symposium on VLSI (GLSVLSI '15). ACM, New York, NY, USA, 15–20. DOI:https://doi.org/10.1145/2742060.2742069

    \bibitem{KELLER}
    C. Keller et al., "Dynamic memory-based physically unclonable function for the generation of unique identifiers and true random numbers," 2014 IEEE International Symposium on Circuits and Systems (ISCAS), 2014, pp. 2740-2743, doi: 10.1109/ISCAS.2014.6865740.

@@

    \bibitem{HASH}
    M. S. Hashemian et al. "A robust authentication methodology using physically unclonable functions in DRAM arrays," 2015 Design, Automation \& Test in Europe Conference \& Exhibition (DATE), 2015, pp. 647-652, doi: 10.7873/DATE.2015.0308.
